---
title: "HEYDAY Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


#SALES FORECASTING

Sales forecasting is an important process for any business owner or proactive sale- force. Forecasting allows for accurate prediction of future sales based on past performance. This makes for a more efficient operation since the business is able to plan future activities. Sales forecasting also plays a key role in the expansion of a business.


###STOCK AND PRODUCT MANAGEMENT

Sales forecasting will allow you to have better control of your inventory. You'll be able to examine trends to determine your peak selling and slow selling periods, so you'll know how much inventory you need to keep on hand during the year. This will help prevent lost sales due to out-of-stock situations, as well as the cost associated with carrying too much inventory.

###CUSTOMER INFORMATION
Sales forecasting allows you to spot trends for your individual customers based on buying patterns. This will help you to spot opportunities to sell beneficial products to customers they hadn't purchased. You'll also be able to identify products customers buy frequently so that you can offer special promotions to increase sales.

###USE FOR SALESPEOPLE
Salespeople can use sales forecasting as a planning tool to maximize commissions and bonuses. By knowing what customers have purchased in the past during specific times of the year, the salespeople can time their sales calls so that they they contact customers when they are ready to buy. This ensures they will be making the best use of their precious time.


###TRACKING
Having a tradition of forecasting sales on a quarterly, semi-annual or annual basis not only helps you plan your business, it also increases your corporate knowledge base. When changes in the economy arise, you can always go back to your previous forecasts for hints on what has and has not worked in the past. This can present a significant advantage over any of your competitors who use the "seat-of-the-pants" method of planning and operating their businesses.

Business owners can use sales forecasting to ensure they are adequately staffed. They can anticipate peak selling periods and hire additional help if necessary. Conversely, if a slow selling period is looming, they will know if they're in a position where they have to reduce worker hours.


##DOLMEN MALL CLIFTON DAILY SALES TREND

```{r}
library(ggplot2)
library(forecast)
data_hdmc<-read.csv("Daily sale Report_final_hdmc_2.csv")

hdmc_ts <- ts(data_hdmc[,2], start = c(2017,5,28), frequency = 365)


autoplot(hdmc_ts,xlab="Date",ylab="Sales")
```


##15 DAY FORECAST FOR DOLMEN MALL CLIFTON


```{r}


model_forecast<-auto.arima(hdmc_ts)
fore<-forecast(model_forecast,15)
fore_df<-data.frame(fore)
final<-setNames(cbind(rownames(fore_df), fore_df, row.names = NULL), 
                c("Date","Forecasted Value", "Low.80", "High.80","Low.95","High.95"))
final_forecast<-final[,-1]
autoplot(fore,xlab="Date",ylab="Sales")

final_forecast

```


##LUCKYONE DAILY SALES TREND

```{r}
library(ggplot2)
library(forecast)

data_lucky<-read.csv("Daily Sale Report_final_hlucky.csv")

hdmc_ts <- ts(data_lucky[,2], start = c(2017,6,9), frequency = 365)


autoplot(hdmc_ts,xlab="Date",ylab="Sales")
```


##15 DAY FORECAST FOR LUCKYONE


```{r}


model_forecast<-splinef(hdmc_ts,h=15)
fore_df<-data.frame(model_forecast)
final<-setNames(cbind(rownames(fore_df), fore_df, row.names = NULL), 
                c("Date","Forecasted Value", "Low.80", "High.80","Low.95","High.95"))
final_forecast<-final[,-1]
autoplot(model_forecast,xlab="Date",ylab="Sales")

final_forecast

```





